<<<<<<< HEAD

# Quantified Self Movement Data Analysis
### Coursera-Practical Machine Learning Project
##### "by Li Qiu"
=======
---
title: Quantified Self Movement Data Analysis
subtitle: Coursera-Practical Machine Learning Project
author: "by Li Qiu"
---

# Quantified Self Movement Data Analysis
### Coursera-Practical Machine Learning Project
##### by Li Qiu

>>>>>>> origin/master

## Introduction  
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.  

In this project, we will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the barbell lifts correctly and incorrectly in 5 different ways.

More information is available from the website here: http://groupware.les.inf.puc-rio.br/har

**Data source:**

*Training data:* https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

*Testing data:* https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv 

## Project Goal
The goal of project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

1. Your submission should consist of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders).
2. You should also apply your machine learning algorithm to the 20 test cases available in the test data above. Please submit your predictions in appropriate format to the programming assignment for automated grading. See the programming assignment for additional details. 


## Data Preprocessing  

```{r,echo=FALSE,eval=FALSE}
#getwd()
#setwd("C:/Users/lqiu/Coursera/Coursera_practical_machine_learning/")
```
Downloading and read the data sources of training data and testing data to data frams.
```{r,message=FALSE,warning=FALSE}
#download data
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./pml-training.csv"
testFile  <- "./pml-testing.csv"
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile)
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile)
}
#read data
trainRaw<-read.csv("./pml-training.csv")
testRaw<-read.csv("./pml-testing.csv")
```
Loading required library for this project
```{r, message=FALSE}
library(caret);library(rpart);library(rpart.plot);library(randomForest);library(corrplot);library(ggplot2)
```
#### Initial check on data

Traning data contain 160 variables and 19622 observations, the testing data contain 160 variables and 20 observations. Also, there are only 406 observations do not have missing value in training data, and all observations have missing value in testing data.
```{r}
dim(trainRaw)
dim(testRaw)
sum(complete.cases(trainRaw))
sum(complete.cases(testRaw))
```
#### Data Cleaning
Setting missing Values ("*NA*") to 0 in both data.
```{r}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0]
```
Remove some variables that do not contribute to the measurements.

```{r}
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
```
#### Data Spliting
Spliting the cleaned traning data to 80% traning dataset and 20% validation dataset.
```{r}
set.seed(1515) #for reproduction purpose
inTrain <- createDataPartition(trainCleaned$classe, p=0.80, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

#### Zero Covariates

Checking for the zero variance variable to collect useful variables for constructing a prediction model. From the following table, there all **nzv** values are *FALSE*. Thus, there is no variable we need to remove.  
```{r,echo=FALSE}
nearZeroVar(trainData,saveMetrics=TRUE)
```

#### Correlation

Following plot shows the correlation between most variables is zero. There are few variables has correlation greater than 0.8, they are same category of exercise, such as *accel_belt_z* and *accel_belt_y*. This problem can be avoid by some algorithms, such as Principal Component Analysis and Random Forest Algorithm.
```{r,echo=FALSE}
corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="color",tl.cex=0.5)
```

## Modeling

Using Random Forest algorithm, because it automatically selects important variables and is robust to correlated covariates and outliers in general. the 3-fold cross validation will be used to apply the algorithm.
```{r}
controlRf <- trainControl(method="cv", 3)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=200)
modelRf
```
Estimate the performace of the model on validation dataset. In the Reference talbe, there are only few cases with prediction error. The model estimated accuracy is about 99.5% and the estimated out-of-sample error is about 0.5%.
```{r}
predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)
#estimated accuracy
postResample(predictRf, testData$classe)
#out of sample error
1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
```
To check the cases where the incorrect prediction. I picked the most 2 important variables to show.
```{r}
#the most 2 important Variable
order(varImp(modelRf)$importance,decreasing=T)[c(1,2)]
#name of the Variable
names(trainData[,c(1,3)])
```
In following plots, there are only few cases(red dots in "Newdata Predictions") with incorrect prediction. Within the compare between the plots, the incorected cases are the place, different classes twisted. 

```{r,echo=FALSE}
p <- qplot(roll_belt,yaw_belt, col=classe,data=trainData,main="Training Data",xlim=c(-30,170),ylim=c(-180,180))
predRight<-predictRf==testData$classe
cpreP<-qplot(roll_belt,yaw_belt,colour=predRight,data=testData,main="Newdata Predictions",xlim=c(-30,170),ylim=c(-180,180))
p;cpreP
```

Showing an example of the decision tree:

```{r,echo=FALSE}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel)
```

## Predict On Test Dataset

Applying model to the original test dataset.

```{r}
result_test <- predict(modelRf, testCleaned[,-length(names(testCleaned))])
result_test
```

```{r,echo=FALSE,eval=FALSE}
pml_write_files = function(x){
  n = length(x)
  path <- getwd()
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=file.path(path, filename),quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(result_test)
```
